---
title: "ADM Quartile Analysis"
author: "Meg McMahon"
date: "3/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(reshape2)
library(broom)
library(knitr)
library(RColorBrewer)
library(ggcorrplot)

```


1) Data cleaning
2) Between age group differences for circadian measures
3) Regression analysis by all subjects, younger adults, and older adults for MemMatch accuracy and reaction time
4) Regression analysis by face-object vs scene-object MemMatch trial for all subjects

#### v2

Note: 01-23 v1 does NOT have the 10 day limit on actigraphy data. V2 takes 10 days prior to scanning session.

```{r settings}

MIN_ACCURACY = 0.55


#Choose combined dataset from CAR_measures.Rmd output

#d1 <- read.csv(file.choose())
d1 <- read.csv("~/Box/ADM/data/combined_data_2019-01-23-v2.csv")

#Choose total sleep time excel sheet
#tst <- readxl::read_xlsx(file.choose())
#tst <- readxl::read_xlsx("~/Box/CogNeuroLab/Aging Decision Making R01/Data/CAR/Average Sleep Time.xlsx")

#Read in all memmatch results
mem <- read.csv("~/Box/ADM/data/summary/memmatch_results.csv")

#d <- merge(d1, tst, by = "record_id")
d <- d1
```


# Data Cleaning

# Overall Demographics
```{r}

d <- subset(d, ! d$record_id == "30412")
d <- subset(d, ! d$record_id == "30242")

d$group[d$age <= 30]<- "YA"
d$group[d$age >= 60] <- "OA"
d$group[d$age >30 && d$age < 60] <- "NULL"

d$sex[substr(d$sex, 0, 1) == "f" | substr(d$sex, 0, 1) == "F"] <- "F"
d$sex[substr(d$sex, 0, 1) == "m" | substr(d$sex, 0, 1) == "M"] <- "M"
d$sex <- factor(d$sex)


d$edu <- substr(d$edu, 0, 2)
d$edu <- str_pad(d$edu, 2, pad = "0")
d$edu <- as.numeric(d$edu)

d$actquotient <- d$actamp/d$actmesor

d %>%
  group_by(group) %>%
  summarize(N = n(), 
            age_mean = mean(age), age_sd = sd(age), 
            Female = sum(sex == "F"), Male = sum(sex == "M"),
            years_edu = mean(edu), edu_sd = sd(edu))
```

Not in dataset: "30003" "30027" "30101" "40511" "40653" "40707" "40733" "40738"


## New Demographics
Excluding participants with accuracy < 0.55

```{r data cleaning}
d$record_id[d$accuracy < 0.55]


d2 <- subset(d, accuracy > 0.55)

d <- d2

d %>%
  group_by(group) %>%
  summarise(n = n(), min_acc = min(accuracy, na.rm = TRUE), max_acc = max(accuracy, na.rm = TRUE), mean_acc=mean(accuracy, na.rm = TRUE), sd_acc=sd(accuracy, na.rm = TRUE), min_rt = min(rt, na.rm = TRUE), max_rt = max(rt, na.rm = TRUE), mean_rt = mean(rt, na.rm = TRUE), sd_rt = sd(rt, na.rm = TRUE))



```

# Demographics
```{r demographics}
d %>%
  group_by(group) %>%
  summarize(N = n(), 
            age_mean = mean(age), age_sd = sd(age), 
            Female = sum(sex == "F"), Male = sum(sex == "M"),
            years_edu = mean(edu), edu_sd = sd(edu))
```

## Accuracy by group
```{r}
#Accuracy by group
t <- t.test(accuracy ~ group, data = d1)
t
d1%>%  
  subset(d1$accuracy > MIN_ACCURACY) %>%
  ggplot(aes(x = group, y = accuracy, color = group)) +
  geom_boxplot(lwd=1.5) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  labs(caption= paste("t =", round(t$statistic, 2), "p = ", round(t$p.value, 3))) +
  theme(plot.caption = element_text(size = 12)) +
  ggsave('~/Desktop/accuracy.png')
  

d %>%
  subset(d$accuracy > MIN_ACCURACY) %>%
  group_by(group) %>%
  summarise(n = n(), min = min(accuracy), median = median(accuracy), mean = mean(accuracy), max = max(accuracy))

```

## Reaction time by group
```{r}
d2.mlt <- melt(d2, id.vars=c("group", "accuracy", "rt"))

#Reaction time by group
t <- t.test(rt ~ group, data = d1)
t
d2.mlt%>%  
  subset(d2$accuracy > MIN_ACCURACY) %>%
  ggplot(aes(x = group, y = rt, color = group)) +
  geom_boxplot(lwd=1.5) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(caption= paste("t =", round(t$statistic, 2),round(t$p.value, 3))) +
  ggsave('~/Desktop/rt.png')

d %>%
  subset(d$accuracy > MIN_ACCURACY) %>%
  group_by(group) %>%
  summarise(n = n(), min = min(rt), median = median(rt), mean = mean(rt), max = max(rt))

```

# Quantile analysis
### Create quantiles

```{r}
# nparACT values
d$ISquant[d$IS > quantile(d$IS)[4]] <- "upper"
d$ISquant[d$IS < quantile(d$IS)[2]] <- "lower"

d$IVquant[d$IV > quantile(d$IV)[4]] <- "upper"
d$IVquant[d$IV < quantile(d$IV)[2]] <- "lower"

d$RAquant[d$RA > quantile(d$RA)[4]] <- "upper"
d$RAquant[d$RA < quantile(d$RA)[2]] <- "lower"

d$M10quant[d$M10 > quantile(d$M10)[4]] <- "upper"
d$M10quant[d$M10 < quantile(d$M10)[2]] <- "lower"

d$L5quant[d$L5 > quantile(d$L5)[4]] <- "upper"
d$L5quant[d$L5 < quantile(d$L5)[2]] <- "lower"

d$ampquant[d$actamp > quantile(d$actamp)[4]] <- "upper"
d$ampquant[d$actamp < quantile(d$actamp)[2]] <- "lower"

d$ampquotquant[d$actquotient > quantile(d$actquotient)[4]] <- "upper"
d$ampquotquant[d$actquotient < quantile(d$actquotient)[2]] <- "lower"

d$upmesquant[d$actupmesor > quantile(d$actupmesor)[4]] <- "upper"
d$upmesquant[d$actupmesor < quantile(d$actupmesor)[2]] <- "lower"

d$downmesquant[d$actdownmesor > quantile(d$actdownmesor)[4]] <- "upper"
d$downmesquant[d$actdownmesor < quantile(d$actdownmesor)[2]] <- "lower"

d$widthquant[d$actbeta > quantile(d$actbeta)[4]] <- "upper"
d$widthquant[d$actbeta < quantile(d$actbeta)[2]] <- "lower"

d$widthratquant[d$actwidthratio > quantile(d$actwidthratio)[4]] <- "upper"
d$widthratquant[d$actwidthratio < quantile(d$actwidthratio)[2]] <- "lower"

d$minquant[d$actmin > quantile(d$actmin)[4]] <- "upper"
d$minquant[d$actmin <= quantile(d$actmin)[2]] <- "lower"


# by group
OA <- subset(d, group == "OA")
YA <- subset(d, group == "YA")

```

### Analysis - Mean Accuracy on MemMatch Task
```{r}

#note: doesn't matter if specify !is.na() result is same
t.test(accuracy ~ ISquant, d[!is.na(d$ISquant),]) # NS p = 0.092, p = 0.092
t.test(accuracy ~ IVquant, d) # NS
t.test(accuracy ~ RAquant, d) # NS
t.test(accuracy ~ M10quant, d) # NS
t.test(accuracy ~ L5quant, d) # NS

t.test(accuracy ~ ISquant, OA) # NS
t.test(accuracy ~ IVquant, OA) # NS
t.test(accuracy ~ RAquant, OA) # NS
t.test(accuracy ~ M10quant, OA) # NS
t.test(accuracy ~ L5quant, OA) # NS

t.test(accuracy ~ ISquant, YA) # NS
t.test(accuracy ~ IVquant, YA) # NS
t.test(accuracy ~ RAquant, YA) # NS
t.test(accuracy ~ M10quant, YA) # NS p = 0.08771
t.test(accuracy ~ L5quant, YA) # NS

# cosinor values
t.test(accuracy ~ ampquant, d) # NS p = 0.063
t.test(accuracy ~ ampquotquant, d) #NS p = 0.073
t.test(accuracy ~ upmesquant, d) # NS
t.test(accuracy ~ downmesquant, d) # NS
t.test(accuracy ~ widthquant, d) # NS
t.test(accuracy ~ widthratquant, d) # NS
t.test(accuracy ~ minquant, d) # p = 0.04577

t.test(accuracy ~ ampquant, OA) # NS p = 0.0834
t.test(accuracy ~ ampquotquant, OA) #NS 
t.test(accuracy ~ upmesquant, OA) # NS
t.test(accuracy ~ downmesquant, OA) # p = 0.0060 **
t.test(accuracy ~ widthquant, OA) # p = 0.01236 **
t.test(accuracy ~ widthratquant, OA) # NS
t.test(accuracy ~ minquant, OA) # NS p = 0.1186

t.test(accuracy ~ ampquant, YA) # NS 
t.test(accuracy ~ ampquotquant, YA) #NS
t.test(accuracy ~ upmesquant, YA) # NS p = 0.06406
t.test(accuracy ~ downmesquant, YA) # NS
t.test(accuracy ~ widthquant, YA) # NS p = 0.08833
t.test(accuracy ~ widthratquant, YA) # NS
t.test(accuracy ~ minquant, YA) # NS

```

### Accuracy Results
#### OA with later down-mesor show greater mean accuracy on MemMatch task
#### OA with smaller width values (alpha) have greater mean accuracy on MemMatch task
#### Participants with smaller width values (alpha) have greater mean accuracy on MemMatch task across age groups

```{r}
#OA with later down-mesor show greater mean accuracy on MemMatch task
ggplot(data = d[!is.na(d$downmesquant),], aes(x = downmesquant, y = accuracy, color = downmesquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by down-mesor quantile") #+
  #ggsave('~/Box/ADM/quantile/downmes_accuracy.png')

ggplot(data = d[!is.na(d$downmesquant),], aes(x = group, y = accuracy, color = downmesquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by down-mesor quantile", caption = "OA, t=-3.3551, p=0.006014") #+
  #ggsave('~/Box/ADM/quantile/downmes_accuracy.png')

#OA with smaller width values (alpha) have greater mean accuracy on MemMatch task
ggplot(data = d[!is.na(d$widthquant),], aes(x = widthquant, y = accuracy, color = widthquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by width (alpha) quantile") 

ggplot(data = d[!is.na(d$widthquant),], aes(x = group, y = accuracy, color = widthquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by width (alpha) quantile", caption = "OA, t=2.725, p=0.01236") #+
  #ggsave('~/Box/ADM/quantile/width_accuracy.png')

#Participants with smaller width values (alpha) have greater mean accuracy on MemMatch task across age groups
ggplot(data = d[!is.na(d$minquant),], aes(x = minquant, y = accuracy, color = minquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by minimum cosinor rhythm quantile", caption = "Across age groups, t=1.8628, p=0.06934")

ggplot(data = d[!is.na(d$minquant),], aes(x = group, y = accuracy, color = minquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by minimum cosinor rhythm quantile") #+
  #ggsave('~/Box/ADM/quantile/width_accuracy.png')
```

### Analysis - RT on correct MemMatch trials
```{r}
t.test(rt ~ ISquant, d[!is.na(d$ISquant),]) # NS
t.test(rt ~ IVquant, d) # NS
t.test(rt ~ RAquant, d) # NS
t.test(rt ~ M10quant, d) # NS
t.test(rt ~ L5quant, d) # NS

t.test(rt ~ ISquant, OA) # NS
t.test(rt ~ IVquant, OA) # NS
t.test(rt ~ RAquant, OA) # NS
t.test(rt ~ M10quant, OA) # NS
t.test(rt ~ L5quant, OA) # NS

t.test(rt ~ ISquant, YA) # NS
t.test(rt ~ IVquant, YA) # NS
t.test(rt ~ RAquant, YA) # NS
t.test(rt ~ M10quant, YA) # NS p = 0.0961
t.test(rt ~ L5quant, YA) # NS

# cosinor values
t.test(rt ~ ampquant, d) # NS p = 0.08134
t.test(rt ~ ampquotquant, d) # t = 2.4816, p = 0.01741 *
t.test(rt ~ upmesquant, d) # NS
t.test(rt ~ downmesquant, d) # t = -3.6015, p = 0.0009332 ***
t.test(rt ~ widthquant, d) # NS p = 0.06611
t.test(rt ~ widthratquant, d) # NS p = 0.07658
t.test(rt ~ minquant, d) # t = -2.2111, p = 0.03298 *

t.test(rt ~ ampquant, OA) # NS
t.test(rt ~ ampquotquant, OA) #NS 
t.test(rt ~ upmesquant, OA) # NS
t.test(rt ~ downmesquant, OA) # t = 2.7261, p = 0.03775 *
t.test(rt ~ widthquant, OA) # NS
t.test(rt ~ widthratquant, OA) # NS
t.test(rt ~ minquant, OA) # NS

t.test(rt ~ ampquant, YA) # NS 
t.test(rt ~ ampquotquant, YA) #NS
t.test(rt ~ upmesquant, YA) # NS
t.test(rt ~ downmesquant, YA) # NS
t.test(rt ~ widthquant, YA) # NS p = 0.08833
t.test(rt ~ widthratquant, YA) # NS
t.test(rt ~ minquant, YA) # NS
```


```{r}
#Participants with greater amplitude quotient values have greater mean accuracy on MemMatch task across age groups
ggplot(data = d[!is.na(d$ampquotquant),], aes(x = ampquotquant, y = accuracy, color = ampquotquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by amplitude quotient quantile", caption = "Across age groups, t=2.4816, p=0.01741") #+
  #ggsave('~/Box/ADM/quantile/ampquot_rt.png')

ggplot(data = d[!is.na(d$ampquotquant),], aes(x = group, y = accuracy, color = ampquotquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by amplitude quotient quantile") #+
  #ggsave('~/Box/ADM/quantile/ampquot_rt.png')

#Participants with later down-mesor have greater mean accuracy on MemMatch task across age groups
ggplot(data = d[!is.na(d$downmesquant),], aes(x = downmesquant, y = accuracy, color = downmesquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by down-mesor quantile", caption = "Across age groups, t=3.6015, p=0.0009332") #+
  #ggsave('~/Box/ADM/quantile/downmesor_rt.png')

ggplot(data = d[!is.na(d$downmesquant),], aes(x = group, y = accuracy, color = downmesquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by down-mesor quantile") #+
  #ggsave('~/Box/ADM/quantile/downmesor_rt.png')

#Participants with lower minimums in their cosinor rhythm have greater mean accuracy on MemMatch task across age groups
ggplot(data = d[!is.na(d$minquant),], aes(x = minquant, y = accuracy, color = minquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by minimum quantile", caption = "Across age groups, t = -2.2111, p = 0.03298") #+
  #ggsave('~/Box/ADM/quantile/downmesor_rt.png')

ggplot(data = d[!is.na(d$minquant),], aes(x = group, y = accuracy, color = minquant)) +
  geom_boxplot(lwd=1.0) + 
  theme_classic() + 
  scale_colour_manual(values = brewer.pal(8, "Paired")[7:8]) +
  theme(plot.caption = element_text(size = 12)) +
  labs(title= "Accuracy differences by minimum quantile") #+
  #ggsave('~/Box/ADM/quantile/downmesor_rt.png')

```

